---
title: "Coupling Modulation"
author: "RRC"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(signal) # library for signal analysis
library(viridis) # colormaps
library(signal, warn.conflicts = F, quietly = T) # signal processing functions
library(oce, warn.conflicts = F, quietly = T) # image plotting functions and nice color maps
library(spectral) 
library(seewave) # huilbert transformation
library(pracma) # findpeaks
library(ggplot2) # polar plot
library(grid) # Arrow vector in polar coordinates
```

```{r functions, include=FALSE, echo=FALSE, message=FALSE}
angle <- function(z) {
    #   Phase angle extraction
    #   angle(H) returns the phase angles, in radians, 
    #   of a matrix withcomplex elements.  
    # The angle function is simply defined as atan2(Im(z), Re(z))
        A <- atan2(Im(z), Re(z))
    return(A)
}

amp <- function(z) {
    #   Amplitud extraction
    #   angle(H) returns the phase angles, in radians, 
    #   of a matrix withcomplex elements.  
    # The angle function is simply defined as atan2(Im(z), Re(z))
        M <- sqrt(Im(z)^2 + Re(z)^2)
    return(M)
}

rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}

nextpot2 <- function (x) 
{
    if (!(is.numeric(x) || is.complex(x))) {
        stop(sprintf("argument %s must be numeric or complex", 
            sQuote("x")))
    }
    if (length(x) == 0) {
        return(numeric(0))
    }
    x[x == 0] <- 1
    return(ceiling(log2(abs(x))))
}

plot.frequency.spectrum <- function(X.k, xlimits=c(0,length(X.k))) {
  plot.data  <- cbind(0:(length(X.k)-1), Mod(X.k))

  # TODO: why this scaling is necessary?
  plot.data[2:length(X.k),2] <- 2*plot.data[2:length(X.k),2] 
  
  plot(plot.data, t="h", lwd=2, main="", 
       xlab="Frequency (Hz)", ylab="Strength", 
       xlim=xlimits, ylim=c(0,max(Mod(plot.data[,2]))))
}

```
  

## 1. Read the data
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# https://hansenjohnson.org/post/spectrograms-in-r/
P <- function(Path){koti=Sys.getenv("KOTI"); return(paste0(koti,Path))}
Data <- read.table(P("git_here/timeSeries/coupling_mod/Basal.atf"),skip = 10,col.names = c("time","mV"))
```

## Spectrogram of the raw time series  
```{r spectro, echo=FALSE, message=FALSE, warning=FALSE}
Ns <- which(Data$time==1) # Seconds to test

# Local Field Potential Signal (LFP)
LFP <- Data$mV[1:Ns]
LFP.t <- Data$time[1:Ns]

# Sampling frecuency (Hz)
Fs <- 5000

# create spectrogram
spec <- specgram(x = LFP,
                n = min(256, length(LFP)), # size of the fourier window
                Fs = Fs,                   # Sample rate, Hz
                # window = window,
                # overlap = overlap
                )

# discard phase information
Phase <- abs(spec$S)

# normalize
Phase <- Phase/max(Phase)

# convert to dB
Phase <- 10*log10(Phase)

# config time axis
Time <- spec$t

# plot spectrogram
imagep(x = Time,
       y = spec$f,
       z = t(Phase),
       col = magma,
       ylab = 'Frequency [Hz]',
       xlab = 'Time [s]',
       drawPalette = T,
       decimate = F
)
```
  
## PCA of the raw time series  
```{r pca, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}

```
  
  
## Filtering of the data
```{r filters, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
# Nyquist ratio
NyFq <- Fs/2

# theta waves are from 25 to 140, $0 of particular interest Hz
# Inrerior gamma limit
Cut20 <- 20/NyFq
# Superior gamma limit
Cut40 <- 40/NyFq

# HIppocampal theta waves are from 6 to 10 Hz
# Inferior theta limit
Cut4 <- 4/NyFq
# Superior theta limit
Cut11 <- 11/NyFq

# Butterworth filter
# Coeficientes Butterworth para gamma 
# [Bg,Ag]=butter(4,[Cut20 Cut40])
gamma.butter <- butter(4, W = c(Cut20, Cut40))

# Coeficientes Butterworth para theta
# [Bt,At]=butter(2,[Cut4 Cut11])
theta.butter <- butter(2, W = c(Cut4, Cut11))

# Forward And Reverse Filter (Zero-phase digital filtering)
Gamma <- filtfilt(gamma.butter, LFP)
# Filtro sin delay theta
Theta <- filtfilt(theta.butter, LFP)

# ----------------------------------------------------------- #
# plot settings
col.theta <- "coral3"
col.gamma <- "cornflowerblue"
time.crop <- LFP.t[1:Ns]
par(mfrow=c(2,1))

# Raw data plot
plot(time.crop, Data$mV[1:Ns], type='l', bty='n', xlab = 'Time (s)', ylab='mV', main='LFP raw data', col='gray45')
abline(h=0, lty=2, col="red4", lwd=1)
# Raw data plot, GAMMA & THETA
plot(time.crop, Data$mV[1:Ns], type='l', bty='n', xlab = 'Time (s)', ylab='mV', main='LFP raw data, Filtered gamma and theta', col='gray45')
abline(h=0, lty=2, col="red4", lwd=1)
lines(time.crop,Gamma[1:Ns], col=col.gamma, lwd=2)
lines(time.crop, Theta[1:Ns], col=col.theta, lwd=2)
legend("bottomleft", c("Gamma", "Theta"),
       col = c(col.gamma, col.theta), fill = NA, border = NA, text.col = c(col.gamma, col.theta), lty=1, box.col = NA, bg = NA)

```
  
  
```{r filters.plot, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
par(mfrow=c(2,1))
# GAMMA over THETA
plot(time.crop, Gamma[1:Ns], type='l', bty='n', xlab='Time (s)', ylab='mV', main="Gamma peaks over theta cycles\n(filter)", col=col.gamma)
abline(h=0, lty=2, col="red4", lwd=1)
lines(time.crop, Theta[1:Ns], col=col.theta, lwd=2)

# SCALED - Gamma peaks over theta
plot(time.crop, Gamma[1:Ns], type='l', bty='n', xlab='Time (s)', ylab='', main="Gamma peaks over theta cycles\n(filtered & scaled)", col=col.gamma, yaxt = "n")
axis(side = 2, col = col.gamma, col.ticks = col.gamma, col.axis=col.gamma)
abline(h=0, lty=2, col="red4", lwd=1)
par(new = TRUE)
plot(time.crop, Theta[1:Ns], type = "l", xaxt = "n", yaxt = "n",bty='n',
     ylab = "", xlab = "", col = col.theta, lty = 1, lwd=1.5)
axis(side = 4, col = col.theta, col.ticks = col.theta, col.axis=col.theta)
legend("topleft", c("Gamma", "Theta"),
       col = c(col.gamma, col.theta), fill = NA, border = NA, text.col = c(col.gamma, col.theta), lty=1, box.col = NA, bg = NA)
```
  
  
```{r hilbert, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
# -------------------------------------------------------------- #
# Gamma processing
require(seewave)
thr <- -0.5 #threshold for peak detection (WHY???)

# 1. Hilbert transformation
hilGamma <- hilbert(Gamma, Fs) # from seewave

# 2. Extraction of the instant phase oin radians
angGamma <- angle(hilGamma)

# 3. Expands the previously calculated angles into a full circle (2*pi)
angGamma <- angGamma-2*pi*floor(angGamma/(2*pi))

# 4. Invertion of the signal in order to find 90 =0 rad
negGam <- -1*angGamma

# 5. Find peaks
subGamma <- findpeaks(as.numeric(negGam), minpeakheight=thr)
TotalGPeaks <- dim(subGamma)[1]

# -------------------------------------------------------------- #
# Theta processing
# 1. Hilbert transformation
Htheta <- hilbert(Theta, Fs) # from seewave

# 2. Extraction of the instant phase oin radians
angTheta <- angle(Htheta)

# 3. Angle (-pi to pi) to Radians (0-2*pi)
#    Expands the previously calculated angles into a full circle (2*pi)
angTheta <- angTheta-2*pi*floor(angTheta/(2*pi))

# 4. Invertion of the signal in order to find 90 =0 rad
negThe <- -1*angTheta

# 5. Find peaks
subTheta <- findpeaks(as.numeric(negThe), minpeakheight=thr)
TotalTPeaks <- dim(subTheta)[1]

# TOTAL GAMMA per THETA
GpeakPerTheta <- TotalGPeaks/TotalTPeaks
print(paste('Gamma peaks per Theta cycle:', GpeakPerTheta))

# -------------------------------------------------------------- #
# QC plots of the Hilbert transform and peak detection
# Hilbert plots
par(mfrow=c(2,2))
plot(Htheta, type='l', col=col.theta, main="Theta Hilbert", bty="n")
plot(hilGamma, type='l', col=col.gamma, main="Gamma Hilbert", bty="n")
 
# Peaks Theta
plot(LFP.t ,negThe, type='l', bty="n", col=col.theta, ylab="Time", main=paste("Total Theta peaks:", TotalTPeaks))
points(LFP.t[subTheta[,2]], subTheta[,1] , pch=19, col="gray45", cex=1)
lines(LFP.t, Re(Htheta), col="gray75")

# Peaks Gammma
plot(LFP.t ,negGam, type='l', bty="n", col=col.gamma, ylab="Time", main=paste("Total Gamma peaks:", TotalGPeaks))
points(LFP.t[subGamma[,2]], subGamma[,1] , pch=19, col="gray45", cex=1)

# Save as matlab variables
# library(R.matlab)
# writeMat("/Users/rcruces/git_here/timeSeries/coupling_mod/LFP_data.mat",
         # LFP=LFP, Time=LFP.t, Rgamma=Gamma, Rtheta=Theta, Rnegamma=negGam, RgamTh=gammaTheta)
```
```{r ReIm, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
# DEFINE as FUNCTION
par(mfrow=c(3,1))
# Real Imaginary
plot(Re(Htheta), type='l', col=col.theta, bty='n', ylab="Amplitude(a.u.)", xlab="Time (sec)", main="Theta Real and Imaginary")
lines(Im(Htheta), col="gray45")

# Phase plot
plot(LFP.t, angTheta, col=col.theta, main="Phase angles in radians &\n Real part ", bty='n', type='l')
par(new = TRUE)
plot(LFP.t, Re(Htheta), type = "l", xaxt = "n", yaxt = "n",bty='n',
     ylab = "", xlab = "", col = "gray45", lty = 1, lwd=1.5)

# Phase angle in radians Theta
plot(LFP.t, Re(Htheta), type='l', bty="n", col=col.theta,  ylab="Amplitude(a.u.)", xlab="Time", main="Amplitud envelope", lwd=3)
lines(LFP.t, abs(Re(Htheta)), col="black", lwd=0.75)
lines(LFP.t, amp(Htheta), col="black", lwd=1.5)
```
  
```{r mvl, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
# mean vector lenght
# antaneous fast oscillation amplitude component in time is represented by the length of the complex vector, whereas the 
# slow oscillation phase of the same time point is represented by the vector angle
# https://neuroimage.usc.edu/brainstorm/Tutorials/TutPac
# -------------------------------------------------------------- #
# MEAN VECTOR LENGHT (MVL)
mvl <- function(At, Tt){
    # INPUTS:
    # At :  At is the amplitude at time point t
    # Tt :  the phase angle at time point t
    #
    # OUTPUT:
    # MVL : mean vector lenght
    # if (length(At) != length(Tt)); not equal length END
    n <- length(At)
    MVL <- abs(sum(At * exp(1i*Tt)) / n)
    return(MVL)
}
mvl(Htheta, angTheta)

# CALCULO DE LA LONGITUD DEL VECTOR
# Theta angles with Gamma peaks
gammaTheta <- angTheta[subGamma[,2]]
# MATLAB
gammaTheta <- c(5.3359, 0.5902, 1.6693, 2.4046, 3.0552, 4.1221, 4.8220, 5.9980, 0.8356, 1.4462, 2.7852, 4.0855, 5.2640, 0.0281, 0.9980, 2.6796, 4.3785, 5.9944, 1.2454, 2.8746, 4.4243, 5.7796, 1.0039, 3.3844, 3.9234, 4.7989, 5.2930, 5.1473, 4.4773)

# calcula el vector unitario medio Euler formula
u <- sum(exp(1i*gammaTheta))
# obtiene el radio resultante (r) normalizado entre 0 y 1
absVecLength=abs(u/TotalGPeaks)
# Angle in pi
AngGammaTheta <- angle(u)
# Angle in radians
angGammaTheta_Rad <- AngGammaTheta-2*pi*floor(AngGammaTheta/(2*pi))
# Angle in degrees
mvl.deg <- rad2deg(angGammaTheta_Rad)

# Plot the distribution and the vector
ggplot(data.frame(GT=rad2deg(gammaTheta)), aes(x=GT)) +
    geom_histogram(color="black", fill="white", bins =24) +
    coord_polar(theta = "x", start = 9/6*pi, direction = -1) +
    ggtitle("Gamma Theta Peaks MVL") +
    scale_x_continuous(breaks = seq(0, 360, by = 60), limits = c(0, 360) ) +
    geom_segment(aes(x = mvl.deg, y = 0, xend = mvl.deg, yend = 4), arrow = arrow(length = unit(0.5, "cm")), color=col.theta)

```


```{r GammaEnv, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
# -------------------------------------------------------------- #
#  GAMA ENVELOPE
envPeak  <-  amp(hilGamma)
# Gamma envelope demean
envDemean <- envPeak-mean(envPeak)

# -------------------------------------------------------------- # 
# FOURIER TRANSFORMATION on BOTH signals
# Frecuencia de sampleo (=n? de puntos/tiempo de registro (s))
Fs <- 5000
# CAlulo del tama?o de la variable
L <- length(envDemean)

# optimizacion del tama?o de las variables para la FFT (el mutiplo de 2 mayor mas cercano)
NFFT <- 2^nextpot2(L)

# transformada ponderada de Fourier para LPF
Y1 <- fft(LFP,NFFT)/L

# WHY ONLY HALF?????
# obtencion de la mitad de los valores absolutos del espectro espejo de LFP
espectro1 <- 2*abs(Y1[1:(NFFT/2+1)])

# transformada ponderada de Fourier para la envDemean
xi <- seq(min(LFP.t), max(LFP.t), length.out = NFFT)
yi <- interp1(LFP.t, as.vector(envDemean), xi, method = "linear")
Y2 <- fft(yi, inverse = TRUE)/L
plot.frequency.spectrum(Y2, xlimits=c(0,40))

# obtenci?n de la mitad de los valores absolutos del espectro espejo de la envDemean
espectro2 <- 2*abs(Y2[1:(NFFT/2+1)])

# filtro de 100 puntos para suavizar ambos espectros
Filter <- rep(1/100,100)
# convolucion de LFP con el filtro
espectro1 <- conv(espectro1, Filter)
# convolucion de la envDemean con el filtro
espectro2 <- conv(espectro2, Filter)
# obtencion del rango de frecuencias a trav?s de la frecuencia de sampleo
Frange <- Fs/2 * seq(0,1,length.out = NFFT/2+1)
```
  
  
# `matlab` code
```{matlab}
# -------------------------------------------------------------- # 
# CALCULO DE LA POTENCIA DE LA ENVOLVENTE
fr5 <- 5;
tmp <- abs(f-fr5);
[~, inx4] <- min(tmp);
fr10 <- 10;
tmp <- abs(f-fr10);
[~, inx10] <- min(tmp);
 
fr20 <- 20;
tmp <- abs(f-fr20);
[~, inx20] <- min(tmp);
fr80 <- 60;
tmp <- abs(f-fr80);
[~, inx80] <- min(tmp);
 
for n <- inx4:inx10
    Area <- (espectro2(n)+espectro2(n+1))/2*(f(n+1)-f(n));
end
GenvPow <- sum(Area);

# -------------------------------------------------------------- # 
# CALCULO DEL PICO MAXIMO DE FRECUENCIA
cerotes <- zeros(length(espectro2(1:inx4-1)),1);# prepara un vector de 0 para aislar theta
newespectro2 <- vertcat(cerotes,espectro2(inx4:inx10));# concatena los ceros al rango de theta
[~,inxMaxE] <- max(newespectro2);
GenvMax <- f(inxMaxE);
 
cerotes <- zeros(length(espectro1(1:inx20-1)),1);# prepara un vector de 0 para aislar gamma
newespectro1 <- vertcat(cerotes,espectro1(inx20:inx80));# concatena los ceros al rango de gamma
[~,inxMaxG] <- max(newespectro1);
GMax <- f(inxMaxG);
 

# -------------------------------------------------------------- # 
# GRAFICAS
subplot(2,1,1)
plot(f,espectro1,f,espectro2)
axis([0 80 0 0.005]) # de 0 a 80 Hz y de 0 a 15e-3 mV2
subplot(2,1,2)
plot(time,gamma,time,envPeak,time,theta)
axis([0 1 -0.35 0.35]) # de 0 a 1 s y de -0.1 a 0.1 mV
shg
 
#  Cross-correlation Gamma Envelope-Theta
[acor,lag]  <-  xcorr(envolvente,theta,'coeff');      # correlacion de las senyales
 
[cor,I]  <-  max((abs(acor)));# coeficiente de correlacion entre ambas senales 
Peaklag  <-  lag(I);#  Posicicion del pico maximo 
timeDiff  <-  Peaklag/Fs; #  lag en s del pico maximo
Peaklagms  <-  timeDiff*1000; #  Peak lag en ms.
 
plot(lag/Fs,acor)
axis([-0.5 0.5 -0.6 0.6])
shg
 
# Autocorrelation of gamma signal and Rhythmicity
[acor,lag]  <-  xcorr(gamma,500);      # correlacion de las senyales
Nacor  <-  acor/max(acor); # Normaliza el maximo de autocorrelation
halfcor  <-  Nacor(fix((length(acor)/2)):end); # mitad de la correlacion
halftime  <-  (lag(fix((length(lag)/2)):end))/Fs;# time of the corresponding lag in sec
 
cor1  <-  halfcor; # redefine la variable halfcor
 
peaks  <-  findpeaks(cor1);    # picos
troughs  <-  findpeaks(-cor1); # valles
 
A  <-  peaks(2) + 1;        # identificacion del primer pico
B  <-  - troughs(1) + 1;  # identificacion del segundo valle
 
Cr <- (A-B)/(A+B);
 
# plot(halftime,halfcor)
# axis([0 0.1 -1  1])
# shg
 
```

```{r}
library()
# Determine the strength of phase‐locking (and its statistical significance) between pairs of quasiperiodic time series (Cazelles & Stone 2003)
# https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.12188
# phase.sync
```


## Measurements to implement
Phase syncrony (PS) https://www.rdocumentation.org/packages/synchrony/versions/0.3.8/topics/phase.sync  
https://github.com/tgouhier/synchrony  
Cross-frequenct phase synchrony (CFS)  
Phase-amplitude coupling (PAC)  
http://visbrain.org/auto_examples/objects/plot_pac_obj.html  
https://github.com/EtienneCmb/tensorpac  
https://pactools.github.io/auto_examples/plot_comodulogram.html  
PAC-t https://github.com/muntam/TF-PAC  
https://www.nature.com/articles/s41598-019-48870-2  
https://srcole.github.io/2016/03/06/pac/  
https://github.com/srcole/qwm/blob/master/pac/Estimating%20PAC.ipynb  

1.Amplitud y frecuencia de l  
a senal envolvente  
1. Acoplamiento de la senal enviolmente cin ka senak principla  
1. Acoplamiento de la senal evolvent eocn los potenciales de accion    

# References   
Cross-Frequency Phase–Phase Coupling between Theta and Gamma Oscillations in the Hippocampus  
Mariano A. Belluscio, Kenji Mizuseki, Robert Schmidt, Richard Kempter and György Buzsáki  
Journal of Neuroscience 11 January 2012, 32 (2) 423-435; DOI: https://doi.org/10.1523/JNEUROSCI.4122-11.2012  
https://www.jneurosci.org/content/32/2/423    
https://towardsdatascience.com/four-ways-to-quantify-synchrony-between-time-series-data-b99136c4a9c9   
  
  
# Further resources  
1. (R for Matlab users)[http://mathesaurus.sourceforge.net/octave-r.html]  
1. mat2r from matconv  
1. (Numpy for R users)[http://mathesaurus.sourceforge.net/r-numpy.html]  
1. (Fourier Transform: A R tutorial)[http://www.di.fc.ul.pt/~jpn/r/fourier/fourier.html]  
